Interactions in the 1-D optical lattice clock

### 2018-07-24 olc_soc/soc_interactions.tex
removed unnecessary text and fixed 'extensive with N' to 'extensive with L'

Putting everything together and letting $G_X\equiv G K_X$, our total
interaction Hamiltonian is
\begin{align}
  H_{\t{int}}
  &= - G_0\v S_0\c\v S_0 - G_1\v S_1\c\v S_1 - 2 G_{0,1}\v S_0\c\v S_1
  \tag*{} \\ &\quad + G_{0,1} \sum
  \p{b_{-q,\bar\sigma}^\dag b_{q\bar\sigma} a_{p+2q,\sigma}^\dag a_{p\sigma}
    - b_{-q,\sigma}^\dag b_{q\bar\sigma} a_{p+2q,\bar\sigma}^\dag a_{p\sigma}}
  \tag*{} \\ &\quad - \sqrt2 G_{0,1} \sum
  \p{d_{\p{p+q}/2}^\dag a_{q,\g} a_{p,\e} + \t{h.c.}},
\end{align}
where
\begin{align}
  d_s^\dag \equiv \f1{\sqrt2}
  \p{b_{s+1/2,\e}^\dag b_{s-1/2,\g}^\dag
    + b_{s-1/2,\e}^\dag b_{s+1/2,\g}^\dag}.
\end{align}
Notes:
\begin{enumerate}[label=(\roman*)]
\item The last set of terms ($\sim d_s^\dag$ and $d_s$) appear only if
  we use driving schemes to match the mean band energies, thereby
  effective setting $\bar E_1=\bar E_0$; these terms can be neglected
  in the absence of such driving.
\item The number of nonvanishing terms in $\v S_n$ is proportional to
  the occupation of band $n$.  If there is a large population of atoms
  in the ground band and a small population of atoms in the excited
  band, then we can safely neglect the $\sim\v S_1\c\v S_1$ term,
  which is doubly small compared to e.g. $\v S_0\c\v S_1$.
\item Similarly, if the ground band is nearly filled with atoms in one
  clock state and the excited band only sparsely populated with atoms
  in the opposite clock state, then we can neglect the terms
  $\sim b_{-q,\bar\sigma}^\dag b_{q\bar\sigma} a_{p+2q,\sigma}^\dag
  a_{p\sigma}$, as $b_{-q,\bar\sigma}^\dag b_{q\bar\sigma}$ manifestly
  addresses a small number of atoms while
  $a_{p+2q,\sigma}^\dag a_{p\sigma}$ is only nonzero for quasi-momenta
  $p$ with $p+2q$ unoccupied.
\item If there is a nonzero population of atoms in the excited band,
  the terms
  $\sim b_{-q,\sigma}^\dag b_{q\bar\sigma} a_{p+2q,\bar\sigma}^\dag
  a_{p\sigma}$ will generally cause delocalization of clock
  excitations across all quasi-momenta (on a time scale
  $\sim1/G_{0,1}$ which is extensive with $N$).
\end{enumerate}

>>>>

Putting everything together and letting $G_X\equiv G K_X$, our total
interaction Hamiltonian is
\begin{align}
  H_{\t{int}}
  &= - G_0\v S_0\c\v S_0 - G_1\v S_1\c\v S_1 - 2 G_{0,1}\v S_0\c\v S_1
  \tag*{} \\ &\quad + G_{0,1} \sum
  \p{b_{-q,\bar\sigma}^\dag b_{q\bar\sigma} a_{p+2q,\sigma}^\dag a_{p\sigma}
    - b_{-q,\sigma}^\dag b_{q\bar\sigma} a_{p+2q,\bar\sigma}^\dag a_{p\sigma}}
  \tag*{} \\ &\quad - \sqrt2 G_{0,1} \sum
  \p{d_{\p{p+q}/2}^\dag a_{q,\g} a_{p,\e} + \t{h.c.}},
\end{align}
where
\begin{align}
  d_s^\dag \equiv \f1{\sqrt2}
  \p{b_{s+1/2,\e}^\dag b_{s-1/2,\g}^\dag
    + b_{s-1/2,\e}^\dag b_{s+1/2,\g}^\dag}.
\end{align}
Notes:
\begin{enumerate}[label=(\roman*)]
\item The last set of terms ($\sim d_s^\dag$ and $d_s$) appear only if
  we use driving schemes to match the mean band energies, thereby
  effectively setting $\bar E_1=\bar E_0$; these terms can be
  neglected in the absence of such driving.
\item If there is a nonzero population of atoms in the excited band,
  the terms
  $\sim b_{-q,\sigma}^\dag b_{q\bar\sigma} a_{p+2q,\bar\sigma}^\dag
  a_{p\sigma}$ will generally cause delocalization of clock
  excitations across all quasi-momenta (on a time scale
  $\sim1/G_{0,1}$ which is extensive with $L$).
\end{enumerate}
